---
title: "p1805_hw1_jc5457"
output: github_document
---

```{r, echo = FALSE, message=FALSE}
library(tidyverse)
library(ggridges)
library(readxl)
library(dplyr)
```

# Problem 1
**Loading the NYC transit dataset**
```{r}
transit = read_csv("data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names()%>%
  select(line, station_name, station_latitude, station_longitude,
         starts_with("route"), entry, vending, entrance_type, ada) %>%
  mutate(entry = as.logical(recode(entry, "YES" = "TRUE", "NO" = "FALSE")),
         vending = as.logical(recode(vending, "YES" = "TRUE", "NO" = "FALSE")))
```

**Describing the NYC transit dataset**
* Here, we have the NYC transit data imported, and keep 19 variables from the original csv file, including:`r names(transit)`

* And then, we transfer the two character variables "entry" and "vending" into logical variables by `recode` function.

**Answering questions**
* The distinct stations in the dataset includes: `r distinct(transit, line, station_name)`
* The number of ADA complaint station is `r sum(transit$ada=="TRUE")`
* The proportion of station entrances / exits without vending allow entrance is `r mean(transit$vending)`

**After reformating data**
```{r}

```
* The number of distinct stations serve the A train is
* The number of the stations that serve the A train, and are ADA compliant is 


# Problem 2
**Read, clean, and combine the Mr./Professor. Trash Wheel sheet**
* give reasonable names
* omit non-dumpster-specific rows
* round sports balls and convert into integer, and add a column indicating source of data
* combine the two data sets
```{r}
MR = read_excel("data/Trash Wheel Collection Data.xlsx", 
                sheet = "Mr. Trash Wheel",
                range = "A2:N550")  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = as.integer(sports_balls), 
         source = "Mr. Trash Wheel")

PROF = read_excel("data/Trash Wheel Collection Data.xlsx", 
                  sheet = "Professor Trash Wheel",
                  range = "A2:M97")  %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(year = as.character(year),
         source = "Professor Trash Wheel")

Trash_Wheel = bind_rows(MR, PROF)
```

* The number of observations in the resulting dataset is `r count(Trash_Wheel)`,
* Key variables include: `r names(Trash_Wheel)`
* The total weight of trash collected by Professor Trash Wheel is 
`r sum(subset(Trash_Wheel, source == "Professor Trash Wheel")$weight_tons)`
* The total number of sports balls collected by Mr. Trash Wheel in 2020 is
`r sum(subset(Trash_Wheel, source == "Mr. Trash Wheel", year = "2020")$sports_balls)`


# Problem 3

date = lubridate::parse_date_time2(date,orders ="mdy", cutoff_2000 = 23)